PAR Digital: Tecnologia em Prol do Ensino Inclusivo

### 2024-07-08 Paper.tex
parte escrita para revisão

Para a integração e entrega contínua (CI/CD), o projeto utiliza ferramentas como 
o GitHub Actions, configurado para executar testes automatizados e builds a cada 
commit, garantindo que a aplicação se mantenha estável e pronta para deployment. 
O deploy é realizado em um servidor localizado na UFMG, permitindo uma escalabilidade
 fácil e eficiente.

{\bfseries Your document will be returned to you for revision if
  modifications are discovered.}

@@

The ``\verb|webmedia|'' document class requires the use of the
``Libertine'' typeface family. Your \TeX\ installation should include
this set of packages. Please do not substitute other typefaces. The
``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
as they will override the built-in typeface families.

\section{Licença}

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

@@

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

\section{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification. Multiple authors may share one affiliation. Authors'
names should not be abbreviated; use full first names wherever
possible. Include authors' e-mail addresses whenever possible.

Grouping authors' names or e-mail addresses, or providing an ``e-mail
alias,'' as shown below, is not acceptable:
\begin{verbatim}
  \author{Brooke Aster, David Mehldau}
  \email{dave,judy,steve@university.edu}
  \email{firstname.lastname@phillips.org}
\end{verbatim}

The \verb|authornote| and \verb|authornotemark| commands allow a note
to apply to multiple authors --- for example, if the first two authors
of an article contributed equally to the work.

If your author list is lengthy, you must define a shortened version of
the list of authors to be used in the page headers, to prevent
overlapping text. The following command should be placed just after
the last \verb|\author{}| definition:
\begin{verbatim}
  \renewcommand{\shortauthors}{McCartney, et al.}
\end{verbatim}
Omitting this command will force the use of a concatenated list of all
of the authors' names, which may result in overlapping text in the
page headers.

The article template's documentation, available at
\url{https://www.acm.org/publications/proceedings-template}, has a
complete explanation of these commands and tips for their effective
use.

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|section|, \verb|subsection|, \verb|subsubsection|, and
\verb|paragraph|. They should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

\section{Tables}

The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{sample-franklin}
  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
    Ewing, Inc. [Public domain], via Wikimedia
    Commons. (\url{https://goo.gl/VLCRBB}).}
  \Description{The 1907 Franklin Model D roadster.}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader. Figure captions go below the figure. Your figures should
{\bfseries also} include a description suitable for screen readers, to
assist the visually-challenged to better understand your work.

Figure captions are placed {\itshape below} the figure.

\subsection{The ``Teaser Figure''}

A ``teaser figure'' is an image, or set of images in one figure, that
are placed after all author and affiliation information, and before
the body of the article, spanning the page. If you wish to have such a
figure in your article, place the command immediately before the
\verb|\maketitle| command:
\begin{verbatim}
  \begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{figure caption}
    \Description{figure description}
  \end{teaserfigure}
\end{verbatim}

\section{Citations and Bibliographies}

The use of \BibTeX\ for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \linebreak \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{bibfile}
\end{verbatim}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the \BibTeX\ file.

Citations and references are numbered by default. A small number of
ACM publications have citations and references formatted in the
``author year'' style; for these exceptions, please include this
command in the {\bfseries preamble} (before \linebreak
``\verb|\begin{document}|'') of your \LaTeX\ source:
\begin{verbatim}
  \citestyle{acmauthoryear}
\end{verbatim}

  Some examples.  A paginated journal article \cite{Abril07}, an
  enumerated journal article \cite{Cohen07}, a reference to an entire
  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
  monograph/whole book in a series (see 2a in spec. document)
  \cite{Harel79}, a divisible-book such as an anthology or compilation
  \cite{Editor00} followed by the same example, however we only output
  the series if the volume number is given \cite{Editor00a} (so
  Editor00a's series should NOT be present since it has no vol. no.),
  a chapter in a divisible book \cite{Spector90}, a chapter in a
  divisible book in a series \cite{Douglass98}, a multi-volume work as
  book \cite{Knuth97}, an article in a proceedings (of a conference,
  symposium, workshop for example) (paginated proceedings article)
  \cite{Andler79}, a proceedings article with all possible elements
  \cite{Smith10}, an example of an enumerated proceedings article
  \cite{VanGundy07}, an informally published work \cite{Harel78}, a
  doctoral dissertation \cite{Clarkson85}, a master's thesis:
  \cite{anisi03}, an online document / world wide web resource
  \cite{Thornburg01, Poker06}, a video game (Case 1)
  \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05} and
  (Case 3) a patent \cite{JoeScientist001}, work accepted for
  publication \cite{rous08}, 'YYYYb'-test for prolific author
  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
  contain 'duplicate' DOI and URLs (some SIAM articles)
  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
  couple of citations with DOIs:
  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

>>>>

Para a integração e entrega contínua (CI/CD), o projeto utiliza
o GitHub Actions, configurado para executar testes automatizados e builds a cada
commit, garantindo que a aplicação se mantenha estável e pronta para deployment.
O deploy é realizado em um servidor localizado na UFMG, permitindo uma escalabilidade
 fácil e eficiente.

% {\bfseries Your document will be returned to you for revision if
%   modifications are discovered.}

@@

O projeto oferece diversas funcionalidades voltadas para o suporte
educacional de estudantes com necessidades especiais. A seguir, são
descritas as principais funcionalidades do sistema, cada uma desempenhando
 um papel crucial na promoção de um ambiente inclusivo e eficaz para o
 aprendizado

\subsection{CRUD de Usuários}
No sistema, os {\bfseries administradores} são responsáveis por gerenciar as
escolas, visualizando relatórios e detalhes como professores e
estratégias.
% \begin{verbatim}
%   \begin{teaserfigure}
%     \includegraphics[width=\textwidth]{sampleteaser}
%     \caption{figure caption}
%     \Description{figure description}
%   \end{teaserfigure}
% \end{verbatim}
% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\linewidth]{sample-franklin}
%   \caption{DANIELE}
%   \Description{tENTANDO COLOCAR UMA IMAGEM AQUI.}
% \end{figure}

As {\bfseries escolas}  criam professores, responsáveis e
desenvolvem o PEI, além de visualizar estratégias e dados dos alunos
e professores.

Os {\bfseries coordenadores}  acompanham o progresso dos alunos
e estratégias, visualizando o andamento dos alunos.

Os {\bfseries professores
regulares}  são os usuários chave, criando e aplicando estratégias, e
adicionando relatórios de progresso.

Os {\bfseries professores AEE}  auxiliam
os regulares, visualizando estratégias do PEI e relatórios, e
participando de fóruns.

Os {\bfseries responsáveis}  adicionam documentos
necessários e acompanham o desenvolvimento dos alunos, visualizando
seus dados e relatórios parciais.

\subsection{Criação do PEI}
A funcionalidade de criação do Plano Educacional Individualizado
(PEI) é uma das mais importantes, permitindo que educadores e 
coordenadores elaborem planos personalizados para atender às 
necessidades específicas de cada aluno.

\subsection{Estratégias do PEI}
Dentro do PEI, os educadores podem definir e documentar diversas 
estratégias pedagógicas adaptadas às necessidades do aluno. Essas 
estratégias são orientações práticas que guiam os professores na 
implementação do plano e ajudam a monitorar o progresso do aluno.

\subsection{Relatório sobre Estratégias}
A funcionalidade de relatórios permite que educadores e coordenadores 
acompanhem o andamento das estratégias definidas no PEI. 
Esses relatórios são fundamentais para avaliar o progresso do aluno, 
identificar áreas que necessitam de ajustes e garantir que as 
estratégias estão sendo eficazmente implementadas.

\subsection{Fórum de Acompanhamento do Aluno}
O fórum de acompanhamento do aluno é uma plataforma colaborativa onde 
professores, coordenadores e outros profissionais podem discutir o 
progresso do aluno, compartilhar insights e propor ajustes no PEI. 
Este fórum promove uma abordagem colaborativa e integrada ao 
acompanhamento educacional.

\subsection{Reutilização de Estratégias}
O sistema possui um banco de estratégias pedagógicas que podem ser 
reutilizadas. Educadores podem consultar esse banco para encontrar 
estratégias que já foram aplicadas com sucesso em situações similares
, facilitando a implementação de práticas comprovadamente eficazes.

\subsection{Compartilhamento de Recursos}
Uma funcionalidade vital do PAR Digital é o compartilhamento de 
recursos, especialmente Tecnologias Assistivas. Esse recurso 
permite que professores e coordenadores acessem ferramentas e 
materiais que podem ser utilizados para apoiar o aprendizado 
dos alunos com necessidades especiais.

\subsection{Perfil do Aluno}
O sistema mantém um perfil detalhado do aluno, que inclui informações
 educacionais e fundamentais. Esse perfil permite que educadores e 
 coordenadores tenham uma visão abrangente das necessidades, 
 capacidades e progressos do aluno, facilitando a personalização 
 do ensino e a monitorização do desenvolvimento educacional.

\section{Licença}
Software acadêmico, licença de tecnologias livres que está
 sendo utilizado de forma de cunho social?

@@

\subsection{Perspectiva Acadêmica}
O PAR Digital representa uma inovação significativa no campo 
da educação inclusiva, oferecendo uma ferramenta tecnológica 
poderosa para a gestão eficiente de Plano Educacional 
Individualizado (PEI). Academicamente, essa ferramenta pode 
servir como um catalisador para a pesquisa e o 
desenvolvimento contínuo de estratégias pedagógicas 
adaptativas.

O acesso a dados detalhados sobre o progresso dos alunos, 
bem como o compartilhamento de estratégias bem-sucedidas 
entre os educadores, pode promover a produção de conhecimento
 acadêmico relevante na área da educação inclusiva.

Além disso, o projeto pode facilitar estudos longitudinais 
sobre o impacto de intervenções específicas no 
desenvolvimento educacional de alunos com deficiência, 
fornecendo insights valiosos para a comunidade acadêmica.

\subsection{Perspectiva Social}
Socialmente, o PAR Digital tem o potencial de promover a 
inclusão e a igualdade de oportunidades na educação. Ao 
oferecer uma plataforma acessível e intuitiva para a 
gestão de PEIs, o sistema capacita os educadores a 
oferecer um suporte mais eficaz aos alunos com deficiência, 
adaptando seus métodos de ensino às necessidades 
individuais de cada estudante. Isso não apenas melhora 
a experiência educacional desses alunos, mas também 
fortalece a coesão social ao reconhecer e valorizar a 
diversidade no ambiente escolar.

Além disso, ao envolver os pais e responsáveis no processo 
educacional por meio do acompanhamento do progresso dos 
alunos, o software promove uma parceria colaborativa entre 
escola e comunidade, fortalecendo os laços sociais e a 
confiança na instituição educacional.
